What's in a name? On the (automated) topic naming of software maintenance activities

### 2010-03-02 icpc/icpc2010.tex
latest changes intro and background

\newtheorem{mydef}{definition}
\usepackage{graphicx}
\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
\usepackage{multirow}%,multicolumn}
\usepackage{slashbox}
\usepackage{wrapfig}
\usepackage{booktabs, colortbl}
\usepackage{amsmath}
%\usepackage{amsmath,amssymb}
%\usepackage{algpseudocode}
\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
%\usepackage{algorithm}%,algorithmic-fix}
\usepackage{url}
\begin{document}

@@

We provide a brief overview of software repository mining and information retrieval.

\subsection{Definitions}
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

% area under the ROC 
% check
\emph{Area of ROC Curve} is the area under the Reciever Operating Characteric curve, we will refer to this as \emph{ROC}. ROC values reflect a score, similar to school letter-grades, reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs. the false positive rate (1 - specificity). A perfect learner has a value of 1.0, reflecting perfect recall and precision.

\subsection{Topic models}
Topic analysis uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). %~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: `list next has iterator add', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this paper a topic
is a set of tokens extracted from commit messages found within a
project's source control system (SCS).

>>>>

\usepackage{comment}
\newtheorem{mydef}{definition}
\usepackage{graphicx}
\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
\usepackage{multirow}%,multicolumn}
\usepackage{slashbox}
\usepackage{wrapfig}
\usepackage{booktabs, colortbl}
\usepackage{amsmath}
%\usepackage{amsmath,amssymb}
%\usepackage{algpseudocode}
\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
%\usepackage{algorithm}%,algorithmic-fix}
\usepackage{url}
\begin{document}

@@

We provide a brief overview of software repository mining and information retrieval.
This work is related to the mining software repositories (MSR)~\cite{msr} research community as it deals expressly with analyzing a project's source control repository, and the messages associated with revisions therein.

\subsection{Definitions}
We will use the following terminology in the rest of the paper.
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

% area under the ROC 
% check
\emph{Area of ROC Curve} is the area under the Reciever Operating Characteric curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs. the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

\subsection{Topic and Concept Extraction}
Topic analysis uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining repositories, in particular analyzing source code, bugs or developer comments.
\begin{comments}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
\end{comments}

Typically a topic analysis tool like LDA will try to find $N$ indepdent word distributions found from within the word distributions of all the messages.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this paper a topic
is a set of tokens extracted from commit messages found within a
project's source control system (SCS).
